---
title: "Recodes and Package Installs"
author: "Ryan Burge"
date: "October 30, 2016"
output: html_document
---


## Downloading and Installing R

You need to have downloaded and installed R and R studio before you begin.

A link to R is [here](https://cran.r-project.org/bin/windows/base/)

A link to R Studio is [here](https://www.rstudio.com/products/rstudio/download/)

## Package Installation

One of the reasons that R is so great is that you can install packages that do really helpful things. The three that you need to complete Assignment 1 are ggplot2, dplyr, and car. You can install them and load them with the following syntax. 

Note: I put a hashtag (#) in front of the install.packages command because I had already installed them on my computer. Just delete those hashtags before you put them in your R console. Make sure to use the library command to actually load them into your R memory. 

```{r message=FALSE, warning=FALSE}
#install.packages("ggplot2")
#install.packages("dplyr")
#install.pacakges("car")
library(ggplot2)
library(dplyr)
library(car)
```

## Load in your data

The below command will load in your data and call it "simon"

The codebook for this data is [here](http://opensiuc.lib.siu.edu/cgi/viewcontent.cgi?article=1010&context=ppi_statepolls)

```{r message=FALSE, warning=FALSE}
simon <- read.csv(url("http://goo.gl/exQA14"))
```

## Recodes
